$Z^\prime$ mass limits and the naturalness of supersymmetry

### 2015-03-31 paper/e6ssm-tuning-paper.tex
Modified plot of benchmark tunings to use dashed, dotted and dash-dotted lines for clarity, and amended text to include this description. Also changed 'it had much lighter gluino mass' to 'it had a much lighter gluino mass'

\caption{\textit{Top panel:} Scatter plot of the fine tuning as a function of the cutoff scale $M_X$ for the four benchmark points given in Table \ref{tab:benchmarks}. \textit{Middle left panel:}  Individual sensitivities for MSSM BM1 plotted against the high scale $M_X$ which give the overall tuning shown by the green line in the top panel. \textit{Middle right panel:} Individual sensitivities for E$_6$SSM BM1 plotted against the high scale $M_X$ which give the overall tuning shown by the blue line in the top panel. \textit{Bottom left panel:} Individual sensitivities for MSSM BM2 plotted against the high scale $M_X$ which give the overall tuning shown by the yellow line in the top panel. \textit{Bottom right panel:} Individual sensitivities for E$_6$SSM BM2 plotted against the high scale $M_X$ which give the overall tuning shown by the red line in the top panel.    }

@@

In the top panel one can see that the MSSM BM1 tuning (green curve) 
steadily climbs as the cutoff scale is increased, as one would expect
when the tuning originates from large soft masses entering from the
RGEs. The panel on the middle left confirms this, showing that the
largest tuning contributions come from $\Delta_{A_t}$ and
$\Delta_{m_{H_u}^2}$ with the former being the larger sensitivity
until $M_X\approx 10^8$ GeV at which point $\Delta_{m_{H_u}^2}$ takes
over, leading to the small kink in overall tuning which can be seen in
the green curve in the top panel. In this case we have chosen a
point with large mixing, which is known to reduce the MSSM tuning.  We
found this does not eliminate the tuning as there is still a strong
sensitivity to $A_t$, but we did find that large mixing lead to less
fine tuning overall for the points we examined.

@@

For E$_6$SSM BM1 the fine tuning is shown by the blue curve in the top
panel of Fig.~\ref{Fig:BMs-varyMX} and the individual sensitivities
are given in the middle right panel. The tuning actually reduces
initially as the cutoff is is increased from $20$ TeV.  This occurs
because the largest sensitivity is initially $\Delta_\lambda$ (shown
in solid blue in the middle right panel).  This contains some terms
proportional to $M_Z^{\prime \, 2}$, which provide the dominant
contribution to this sensitivity at very low $M_X$.  However as $M_X$
is increased contributions from the soft masses become more important
and these actually start to cancel the large contribution to
$\Delta_\lambda$ coming from $M_{Z^\prime}$ until $\Delta_\lambda$
passes through zero.  At the same time though these large soft masses
also cause other sensitivities to grow, in particular $\Delta_{M_3}$.
The fine tuning rises with $M_X$ once $M_X \gtrsim 10^5-10^6$ GeV, but
remains lower than that of the other points, until $M_X \approx 10^8$
GeV.  Eventually the $\Delta_{M_3}$ sensitivity leads to this point
being the most fine tuned of the four shown in
Fig.~\ref{Fig:BMs-varyMX}.

@@

By contrast the tuning for E$_6$SSM BM2 is very different, as is shown
in red in the top panel, with the individual sensitivities given in the
bottom right panel. This point was chosen as it had much lighter
gluino mass that is just above the experimental limit of $1.4$ TeV
\cite{Aad:2014lra}.  At $20$ TeV this benchmark is not amongst the
lowest tuned points, since at that scale the tree level tuning from
$M_{Z^\prime}$ dominates.  However the reduction in $M_3$ means that
$\Delta_{M_3}$ is substantially lower and only becomes the dominant
tuning at a much larger scale of $M_X\gtrsim 10^{12}-10^{13}$ GeV,
giving a tuning at $10^{16}$ GeV of $\approx 546$, which is far below
that of the other three benchmark points.

@@

MSSM benchmark BM2 (yellow in top panel, individual sensitivities in bottom left panel) is designed to be similar to E$_6$SSM BM2, for a reasonable comparison.  However from the individual sensitivities one can see that the behaviour is quite similar to MSSM BM1, though in this case the $\Delta_{m_{H_u}^2}$ becomes the largest tuning at a higher $M_X$ and does not reach such large values, since more of the tuning is from the mixing in this case.

>>>>

\caption{\textit{Top panel:} Scatter plot of the fine tuning as a function of the cutoff scale $M_X$ for the four benchmark points given in Table \ref{tab:benchmarks}. \textit{Middle left panel:}  Individual sensitivities for MSSM BM1 plotted against the high scale $M_X$ which give the overall tuning shown by the green (dotted) line in the top panel. \textit{Middle right panel:} Individual sensitivities for E$_6$SSM BM1 plotted against the high scale $M_X$ which give the overall tuning shown by the blue (solid) line in the top panel. \textit{Bottom left panel:} Individual sensitivities for MSSM BM2 plotted against the high scale $M_X$ which give the overall tuning shown by the yellow (dash-dotted) line in the top panel. \textit{Bottom right panel:} Individual sensitivities for E$_6$SSM BM2 plotted against the high scale $M_X$ which give the overall tuning shown by the red (dashed) line in the top panel.    }

@@

In the top panel one can see that the MSSM BM1 tuning (dotted green curve) 
steadily climbs as the cutoff scale is increased, as one would expect
when the tuning originates from large soft masses entering from the
RGEs. The panel on the middle left confirms this, showing that the
largest tuning contributions come from $\Delta_{A_t}$ and
$\Delta_{m_{H_u}^2}$ with the former being the larger sensitivity
until $M_X\approx 10^8$ GeV at which point $\Delta_{m_{H_u}^2}$ takes
over, leading to the small kink in overall tuning which can be seen in
the green curve in the top panel. In this case we have chosen a
point with large mixing, which is known to reduce the MSSM tuning.  We
found this does not eliminate the tuning as there is still a strong
sensitivity to $A_t$, but we did find that large mixing lead to less
fine tuning overall for the points we examined.

@@

For E$_6$SSM BM1 the fine tuning is shown by the solid blue curve in the top
panel of Fig.~\ref{Fig:BMs-varyMX} and the individual sensitivities
are given in the middle right panel. The tuning actually reduces
initially as the cutoff is is increased from $20$ TeV.  This occurs
because the largest sensitivity is initially $\Delta_\lambda$ (shown
in solid blue in the middle right panel).  This contains some terms
proportional to $M_Z^{\prime \, 2}$, which provide the dominant
contribution to this sensitivity at very low $M_X$.  However as $M_X$
is increased contributions from the soft masses become more important
and these actually start to cancel the large contribution to
$\Delta_\lambda$ coming from $M_{Z^\prime}$ until $\Delta_\lambda$
passes through zero.  At the same time though these large soft masses
also cause other sensitivities to grow, in particular $\Delta_{M_3}$.
The fine tuning rises with $M_X$ once $M_X \gtrsim 10^5-10^6$ GeV, but
remains lower than that of the other points, until $M_X \approx 10^8$
GeV.  Eventually the $\Delta_{M_3}$ sensitivity leads to this point
being the most fine tuned of the four shown in
Fig.~\ref{Fig:BMs-varyMX}.

@@

By contrast the tuning for E$_6$SSM BM2 is very different, as is shown
in red (dashed) in the top panel, with the individual sensitivities given in the
bottom right panel. This point was chosen as it had a much lighter
gluino mass that is just above the experimental limit of $1.4$ TeV
\cite{Aad:2014lra}.  At $20$ TeV this benchmark is not amongst the
lowest tuned points, since at that scale the tree level tuning from
$M_{Z^\prime}$ dominates.  However the reduction in $M_3$ means that
$\Delta_{M_3}$ is substantially lower and only becomes the dominant
tuning at a much larger scale of $M_X\gtrsim 10^{12}-10^{13}$ GeV,
giving a tuning at $10^{16}$ GeV of $\approx 546$, which is far below
that of the other three benchmark points.

@@

MSSM benchmark BM2 (yellow, dash-dotted in top panel, individual sensitivities in bottom left panel) is designed to be similar to E$_6$SSM BM2, for a reasonable comparison.  However from the individual sensitivities one can see that the behaviour is quite similar to MSSM BM1, though in this case the $\Delta_{m_{H_u}^2}$ becomes the largest tuning at a higher $M_X$ and does not reach such large values, since more of the tuning is from the mixing in this case.
